% !TEX root = main.tex

\section{Glossary}
\label{se:glossary}

\noindent {\bf Complete analysis.} Analysis that guarantees no false positives, i.e., all reported property violations are true.

\smallskip\noindent {\bf Concrete execution.} An execution of a program using concrete inputs in a real-world environment.

\smallskip\noindent {\bf Concolic execution.} \ldots

\smallskip\noindent {\bf Control flow graph (CFG).} Representation of a program that uses nodes to model instructions and edges to model the control flow between them.

\smallskip\noindent {\bf Control flow path.} Path in the control flow graph of a program. Represents the sequence of instructions executed by the program for a given concrete input.

\smallskip\noindent {\bf Decidable analysis} \ldots

\smallskip\noindent {\bf Model checker.} Given a model of a system, a model checker exhaustively and automatically checks whether the model meets a given specification.

\smallskip\noindent {\bf Symbolic execution.} \ldots

\smallskip\noindent {\bf Symbolic store.} \ldots

\smallskip\noindent {\bf SMT solver.} A Satisfiability Modulo Theories (SMT) instance is a formula in first-order logic, where some function and predicate symbols have additional interpretations, and SMT is the problem of determining whether such a formula is satisfiable. A SMT solver is a tool able to reason over SMT formulas.

\smallskip\noindent {\bf Sound analysis.} Analysis that guarantees no false negatives, i.e., if there is a property violation, then it is reported.

\smallskip\noindent {\bf Path constraints.} \ldots